---
title: "DSC1105_FA1"
author: "Leiana Mari D. Aquino"
date: "2025-01-27"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# EDA_1 : The CyTOF One Experiment Data

## Libraries and Database Loading

### Environment Set-up

The following code is written in R. The packages are pre-written functionality to perform necessary tasks.

```{r echo = FALSE, message = FALSE, warning = FALSE}

# install necessary R libraries
library(tidyverse) # data manipulation
library(dplyr)
library(ggpubr) # publication ready plots
library(ggplot2) # data visualization
library(knitr) # data visualization
library(kableExtra)
```

### Import the data

```{r echo = FALSE, message = FALSE, warning = FALSE}

# load the data using read_csv()
data <- read_csv("cytof_one_experiment.csv")

# view the first six data frame records
six_data_head <- head(data,6)

# print the sum_data_head
kbl(six_data_head) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```

### Summary of the data

```{r echo = FALSE, message = FALSE, warning = FALSE}

# summarize the data
sum_data <- summary(data)

# print the sum_data
kbl(sum_data) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```

## The NKp30 Data Distribution Plots

From the `cytof_one_experiment.csv` data, we will extract the `NKp30` column to make two plots showing its distribution of the values.

### Empirical Cumulative Distribution Function (ECDF)

```{r echo = FALSE, message = FALSE, warning = FALSE}

# extract/grab the column containing the data
NKp30 <- data$NKp30

# ECDF plot using ggplot2 package
cytof.gg <- ggplot(data, aes(x = NKp30)) + 
  stat_ecdf(geom="step", color="#1f6e34") +
  labs(title = "ECDF of NKp30 from CyTOF data", x = "NKp30", y = "ECDF")
cytof.gg

```

| **Interpretation**: This plot show the ECDF of the values in NKP30. A concentration of values from ranges 2 to 4 rises steeply but as it enters the range of 6, the curve became relatively flat indicating fewer data points or gaps in the distribution.

------------------------------------------------------------------------

### Density Estimate

```{r echo = FALSE, message = FALSE, warning = FALSE}

# extract/grab the column containing the data
NKp30 <- data$NKp30

# density estimate plot using ggplot2 package
density_est.gg <- ggplot(data, aes(x = NKp30)) + 
  geom_density(adjust = .5, fill="#77bd89", color="#1f6e34", alpha=0.8) +
  geom_rug(aes(y = 0), sides = "b", position = position_jitter(height = 0)) +
  labs(title = "Density Estimate of NKp30 from CyTOF data", x = "NKp30", y = "Density")
density_est.gg

```

| **Interpretation**: This plot provides a smoothed estimate of the distribution of NKp30. The highest density is around NKp30 = 4, thus, suggesting that most of the values are concentrated in the middle range, with fewer values at the extremes.

## Q-Q Plots

```{r echo = FALSE, message = FALSE, warning = FALSE}

# qq plot using ggplot2 package
qq_plot.gg <- ggplot(data) +
  stat_qq(aes(sample = NKp30), color = "#316da6") +  # qq plot for column1
  stat_qq(aes(sample = KIR3DL1), color = "#1f6e34") + # qq plot for column2
  stat_qq_line(aes(sample = NKp30), color = "#e87c45") +  # reference line for column1
  labs(title = "Q-Q Plot Comparing NKp30 and KIR3DL1",
       x = "NKp30",
       y = "KIR3DL1") 
qq_plot.gg

```

| **Interpretation**: This QQ plot compares the quantiles NKp30 and KIR3DL1 in order to assess whether the two data sets have similar distributions. Based on the provided plot, only the NKp30 (blue-colored) plot lies close to the reference line while the KIR3DL1 (green-colored) plot deviates from the line. These indicates that there are differences in the distributions of the points, for instance, NKp30 is curved pattern while KIR3DL1 is curved in a S-shape.
